---
title: "`r params$title`"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
params:
  title: "Unnamed Report"
  reportID: FALSE
  orgImage: FALSE
  iePar: FALSE
  imwGrid: FALSE
  analyteTable: FALSE
  bgPar: FALSE
  bgPlots: FALSE
  bgTable: FALSE
  ppPlot: FALSE
  ppPar: FALSE
  sessionInfo: FALSE
---
```{r echo=FALSE}
box::use(
  app/logic/gridding[render_grid],
  app/logic/labels_io[create_labels_df],
  app/logic/const[LETTERS_EXT],
  EBImage[display],
)
```

`r if (params$reportID) paste0("Report ID: ", array_data$date, "-", sprintf("%03d", array_data$id))`

`r if (params$orgImage) "## Original Image"`
```{r echo=FALSE, eval=params$orgImage}
display(array_data$imageFinal)
```

`r if (params$iePar) "## Image Editor Parameters"`
```{r echo=FALSE, eval=params$iePar}
ie_params <- data.frame(
  parameters = c("Image Name",
                 "Grid size",
                 "Cell size",
                 "Grid mode"),
  values = c(array_data$imageName,
             paste(array_data$roi$ncols, "x", array_data$roi$nrows),
             paste0(round(array_data$roi$cell_w, 2), "px", " x ",
                    round(array_data$roi$cell_h, 2), "px"),
             if (array_data$roi$mode == "rect") "Rectangular" else "Parallelogram")
)
ie_params
```

`r if (params$imwGrid) "## Image with Grid"`
```{r echo=FALSE, eval=params$imwGrid}
display(array_data$roi$image, method = "raster", margin = c(0, 32, 0, 0))
render_grid(array_data$roi$image, array_data$roi$nrows, array_data$roi$ncols, NULL,
            array_data$roi$mode, analyte_grid = array_data$roi$grid, show_an = FALSE,
            analyte_names = array_data$analytes, show_checkerboard = TRUE,
            selected_analyte = FALSE)
```

`r if (params$analyteTable & !is.null(array_data$analytes)) "## Analyte Table"`
```{r echo=FALSE, eval=(params$analyteTable & !is.null(array_data$analytes))}
cat(paste0("Labels filename: ", array_data$label_filename))
create_labels_df(array_data$roi$grid, array_data$analytes)
```

`r if (params$bgPlots & !is.null(array_data$thresh_data)) "## Background Correction"`
```{r echo=FALSE, fig.show='hold', out.width='50%', eval=(params$bgPlots & !is.null(array_data$thresh_data))}
display(array_data$thresh_data$image_above_bg, method = "raster", margin = c(0, 30, 0, 0))
render_grid(array_data$thresh_data$image_above_bg, array_data$roi$nrows, array_data$roi$ncols,
            NULL, array_data$roi$mode, analyte_grid = NULL, analyte_names = NULL,
            show_checkerboard = TRUE)

display(array_data$thresh_data$image_after_bc, method = "raster", margin = c(0, 30, 0, 0))
render_grid(array_data$thresh_data$image_above_bg, array_data$roi$nrows, array_data$roi$ncols,
            NULL, array_data$roi$mode, analyte_grid = NULL, analyte_names = NULL,
            show_checkerboard = TRUE)
```

`r if (params$bgTable & !is.null(array_data$thresh_data)) "## Background Correction Table"`
```{r echo=FALSE, eval=(params$bgTable & !is.null(array_data$thresh_data))}
bc_df <- data.frame(Cell = as.vector(sapply(LETTERS_EXT[seq_len(array_data$roi$ncols)],
                                               function(x) {
                                                 paste0(x, seq_len(array_data$roi$nrows))
                                                 }
                                               )),
                       Threhold = as.vector(array_data$thresh_data$threshold),
                       Mean = as.vector(array_data$thresh_data$mean_intensities),
                       Median = as.vector(array_data$thresh_data$median_intensities)
)
bc_df
```

`r if (params$bgPar & !is.null(array_data$thresh_data)) "## Background Correction Parameters"`
```{r echo=FALSE, eval=(params$bgPar & !is.null(array_data$thresh_data))}
bc_params <- data.frame(
  parameters = c(
    "Color Image",
    if (array_data$thresh_data$params$colorImage == "Yes") "Channel",
    "Spot Type",
    "Threshold Method",
    if (array_data$thresh_data$params$thresh == "Otsu") "Otsu Mode",
    if (array_data$thresh_data$params$thresh == "Quantile") "Quantile Probability"
  ),
  values = c(
    array_data$thresh_data$params$colorImage,
    if (array_data$thresh_data$params$colorImage == "Yes") array_data$thresh_data$params$channel,
    array_data$thresh_data$params$spotType,
    array_data$thresh_data$params$thresh,
    if (array_data$thresh_data$params$thresh == "Otsu") array_data$thresh_data$params$otsuMode,
    if (array_data$thresh_data$params$thresh == "Quantile") array_data$thresh_data$params$quantile1
  )
)
bc_params
```

`r if (params$ppPlot & !is.null(array_data$thresh_data)) "## Post-Processing Plot"`
```{r echo=FALSE, eval=(params$ppPlot & !is.null(array_data$thresh_data))}
display(array_data$roi$image, method = "raster", margin = c(0, 30, 0, 0))
render_grid(array_data$roi$image, array_data$roi$nrows, array_data$roi$ncols, NULL,
  array_data$roi$mode,
  analyte_grid = array_data$roi$grid,
  analyte_names = array_data$analytes, show_checkerboard = TRUE,
  show_fp = TRUE,
  fp_grid = array_data$ppParams$fp_grid,
  show_id = FALSE,
  id_list = id_list
)
```

`r if (params$ppPar & !is.null(array_data$thresh_data)) "## Post-Processing Parameters"`
```{r echo=FALSE, eval=(params$ppPar & !is.null(array_data$thresh_data))}
pp_params <- data.frame(
  parameters = c(
    "Empty IDs",
    "Positive Control IDs",
    "False Positive Threshold"
  ),
  values = c(
    if (is.null(array_data$ppParams$emptyCells)) "-"
    else paste(array_data$ppParams$emptyCells, collapse = ", "),
    if (is.null(array_data$ppParams$posControl)) "-"
    else paste(array_data$ppParams$posControl, collapse = ", "),
    array_data$ppParams$fp_thresh
  )
)
pp_params
```

`r if (params$sessionInfo) "## Session Info"`
```{r echo=FALSE, warning=FALSE, eval=params$sessionInfo}
sessionInfo()
```

